---
title: "Coverage: Replicate Figures and Tables"
---

Summary: Code to replicate figures and tables in "Berkeley Unified Numident Mortality Database: Public Administrative Records for Individual-Level Mortality Research" section "The Structure and Content of the NARA Numident Records" 

To replicate, move a copy of the BUNMD V2 file into the top-level "data" directory. 

## Setup 

Library packages and source functions 

```{r}
## Library packages 
library(tidyverse)
library(data.table)
library(memisc)
library(here)
library(HMDHFDplus)
library(gridExtra)
library(ggpubr)
library(cowplot)
library(gt)
library(fixest)
library(stargazer)
library(choroplethrZip)
library(choroplethr)
library(sf)
library(broom)
library(lubridate)

## source functions 
source(here("code/helper_functions.R"))

## Read in BUNMD file
bunmd <- fread(here("data/bunmd_v2.csv"))
hmd <- fread(here("data/hmd_deaths_1x1_usa.csv"))
```


```{r}
## number of apps per person
number_apps <- bunmd %>% 
  mutate(number_apps = case_when(
    number_apps >= 3 ~ "3+",
    TRUE ~ as.character(number_apps))) %>% 
    count(number_apps) 

## number of claims per person
number_claims <- bunmd %>% 
  mutate(number_claims = case_when(
    number_claims >= 3 ~ "3+",
    TRUE ~ as.character(number_claims))) %>% 
    count(number_claims) 

## plot claims 
claims_plot <- ggplot(data = number_claims) + 
  geom_col(aes(x = number_claims, y = n), colour="black", fill="grey") + 
  theme(legend.position="bottom") +
  theme_cowplot() + 
  labs(title = "Claims",
       x = "Number of Claim Entries", 
       y = "Count") + 
    scale_y_continuous(labels = scales::comma, limits = c(0, 35000000))
  
## plot apps 
apps_plot <- ggplot(data = number_apps) + 
  geom_col(aes(x = number_apps, y = n), colour="black", fill="grey") + 
  theme(legend.position="bottom") +
  theme_cowplot() + 
  labs(title = "Applications",
       x = "Number of Application Entries", 
       y = "Count") + 
  scale_y_continuous(labels = scales::comma, limits = c(0, 35000000))


## combine plots 
apps_claims_per_person <- cowplot::plot_grid(apps_plot, claims_plot, labels = "auto")

## save plots 
ggsave(plot = apps_claims_per_person, filename = here("figs/apps_claims_per_person.pdf"), width = 10, height = 4)
```


## Compare BUNMD and HMD Death Coverage 

```{r}
## Get HMD deaths from website
hmd <- hmd %>%
  mutate(linking_key = paste(Year, Age, sep = "_" ))

## Tabulate deaths in BUNMD for 65+
bunmd.deaths.tabulated <- bunmd %>% 
  filter(death_age >= 65) %>% 
  group_by(Year = dyear) %>% 
  summarize(deaths = n()) %>% 
  mutate(source = "BUNMD")

## Tabulate deaths in BUNMD for 65+
bunmd.deaths.tabulated_complete <- bunmd %>% 
  filter(!is.na(ccweight)) %>% 
  filter(death_age >= 65) %>% 
  group_by(Year = dyear) %>% 
  summarize(deaths = n()) %>% 
  mutate(source = "BUNMD Complete Case")

## Tabulate deaths in HMD for 65+ 
hmd.deaths.tabulated <- hmd %>%
  filter(Age >= 65) %>% 
  group_by(Year) %>% 
  summarize(deaths = sum(Total)) %>% 
  mutate(source = "Human Mortality Database")

## Combine into one data frame 
data.for.plot <- bunmd.deaths.tabulated_complete %>% 
  bind_rows(bunmd.deaths.tabulated) %>% 
  bind_rows(hmd.deaths.tabulated) %>% 
  filter(Year > 1970 & Year < 2006)

## Create death coverage plot
bunmd_coverage <- ggplot(data.for.plot) + 
  geom_line((aes(x = Year, y = deaths, linetype = source)), size = 1) + 
  labs(x = "Year", 
       y = "Count") + 
  scale_y_continuous(labels = scales::comma) + 
  scale_x_continuous(breaks = scales::pretty_breaks(n=10)) + 
  scale_linetype_manual(values=c(2,3,1)) + 
  theme_cowplot() + 
  theme(legend.position="bottom", legend.title = element_blank()) +
  theme(legend.key.width=unit(1.5, "cm"))

ggsave(plot = bunmd_coverage, filename = here("figs/bunmd_coverage.pdf"), width = 8, height = 5)
```

## Lexis plots  

Create lexis plots comparing HMD and BUNDM death counts. 

```{r}
## Tabulate BUNMD 
bunmd.tabulation <- bunmd %>% 
  filter(death_age %in% 0:100) %>% 
  group_by(Year = dyear, Age = death_age) %>% 
  filter(Year %in% c(1940:2010)) %>% 
  summarize(deaths = n()) 

## HMD tabulation
hmd.tabulation <- hmd %>% 
  filter(Age %in% 0:100) %>% 
  group_by(Year, Age) %>% 
  filter(Year %in% c(1940:2010)) %>% 
  summarize(deaths = sum(Total)) %>% 
  mutate(source = "HMD")

## Prep data for Lexis Function
lexis.data <- bunmd.tabulation %>% 
  inner_join(hmd.tabulation, by = c("Year", "Age" )) %>%
  mutate(coverage = deaths.x/deaths.y)

## Call Lexis Function
lexis_coverage <- lexis(data = lexis.data, year_begin = 1940, age_begin = 0, fill_column = coverage, breaks = c(0, 0.25, .5, .75, .9, .95, 5),
                        labels = c( "0 to 25%",
                                    "25% to 50%",
                                    "50% to 75%",
                                    "75% to 90%",
                                    "90% to 95%",
                                    ">95% Death Coverage")) +
  labs(
    x = "",
    y = "Average Rating (0-10)",
    title = "Coverage",
    subtitle = ""
  ) + theme(legend.position="none")

## Prep data for Lexis Function
lexis.data.filter <- lexis.data %>% 
  filter(Year > 1985 & Age > 65)

## Call Lexis Function
lexis_coverage_all <- lexis(data = lexis.data.filter, year_begin = 1985, age_begin = 65, fill_column = coverage, breaks = c(0, 0.25, .5, .75, .9, .95, 5),
                            labels = c( "0%to 25%",
                                        "25% to 50%",
                                        "50% to 75%",
                                        "75% to 90%",
                                        "90% to 95%",
                                        ">95% Death Coverage")) +
  labs(
    x = "",
    y = "Average Rating (0-10)",
    title = "High Coverage Subsample",
    subtitle = ""
  ) + theme(legend.position="none")

legend_b <- get_legend(
    lexis_coverage + 
        guides(color = guide_legend(nrow = 1)) +
        theme(legend.position = "bottom")
)

## combine lexis plots 
bunmd_lexis <- cowplot::plot_grid(lexis_coverage, lexis_coverage_all, labels = "auto")

# add the legend underneath the row we made earlier. Give it 10%
# of the height of one plot (via rel_heights).
bunmd_lexis <- plot_grid(bunmd_lexis, legend_b, ncol = 1, rel_heights = c(1, .1))

ggsave(plot = bunmd_lexis, filename = here("figs/bunmd_lexis_coverage.pdf"), width = 12, height = 7)


## Call Lexis Function
lexis_coverage_single <- lexis(data = lexis.data, year_begin = 1940, age_begin = 0, fill_column = coverage, breaks = c(0, 0.25, .5, .75, .9, .95, 5),
                        labels = c( "0 to 25%",
                                    "25% to 50%",
                                    "50% to 75%",
                                    "75% to 90%",
                                    "90% to 95%",
                                    ">95% Death Coverage")) +
  theme(legend.position = "bottom") + 
  theme(axis.title.x = element_text(vjust=-1)) 

ggsave(plot = lexis_coverage_single, filename = here("figs/bunmd_lexis_coverage.pdf"), width = 6, height = 7)
```




