---
title: "Replicate Descriptive BUNMD Figures "
author: Casey Breen
---

Summary: Code to replicate figures and tables in "Berkeley Unified Numident Mortality Database: Public Administrative Records for Individual-Level Mortality Research" section "The Structure and Content of the NARA Numident Records" 

This code constructs the following Figures: 

* Figure 1
* Figure 2 
* Figure 3
* Figure A-2
* Figure A-3 

## Setup 

Library packages and source functions 

```{r}
## Library packages 
library(tidyverse)
library(data.table)
library(memisc)
library(here)
library(HMDHFDplus)
library(gridExtra)
library(ggpubr)
library(cowplot)
library(gt)
library(fixest)
library(stargazer)
library(choroplethrZip)
library(choroplethr)
library(sf)
library(broom)
library(lubridate)

## source functions 
source(here("code/helper_functions.R"))

## Read in BUNMD file
bunmd <- fread(here("data/bunmd_v2.csv"))
hmd <- fread(here("data/hmd_deaths_1x1_usa.csv"))
```

# Figure 1 — Compare BUNMD and HMD Death Coverage 


```{r}
## Get HMD deaths from website
hmd <- hmd %>%
  mutate(linking_key = paste(Year, Age, sep = "_" ))

## Tabulate deaths in BUNMD for 65+
bunmd.deaths.tabulated <- bunmd %>% 
  filter(death_age >= 65) %>% 
  group_by(Year = dyear) %>% 
  summarize(deaths = n()) %>% 
  mutate(source = "BUNMD")

## Tabulate deaths in BUNMD for 65+
bunmd.deaths.tabulated_complete <- bunmd %>% 
  filter(!is.na(ccweight)) %>% 
  filter(death_age >= 65) %>% 
  group_by(Year = dyear) %>% 
  summarize(deaths = n()) %>% 
  mutate(source = "BUNMD Complete Case")

## Tabulate deaths in HMD for 65+ 
hmd.deaths.tabulated <- hmd %>%
  filter(Age >= 65) %>% 
  group_by(Year) %>% 
  summarize(deaths = sum(Total)) %>% 
  mutate(source = "Human Mortality Database")

## Combine into one data frame 
data.for.plot <- bunmd.deaths.tabulated_complete %>% 
  bind_rows(bunmd.deaths.tabulated) %>% 
  bind_rows(hmd.deaths.tabulated) %>% 
  filter(Year > 1970 & Year < 2006)

## Create death coverage plot
bunmd_coverage <- ggplot(data.for.plot) + 
  geom_line((aes(x = Year, y = deaths, linetype = source)), size = 1) + 
  labs(x = "Year", 
       y = "Count") + 
  scale_y_continuous(labels = scales::comma) + 
  scale_x_continuous(breaks = scales::pretty_breaks(n=10)) + 
  scale_linetype_manual(values=c(2,3,1)) + 
  theme_cowplot() + 
  theme(legend.position="bottom", legend.title = element_blank()) +
  theme(legend.key.width=unit(1.5, "cm"))

## save plot 
ggsave(plot = bunmd_coverage, filename = here("figs/01_bunmd_coverage.pdf"), width = 8, height = 5)
```

# Figure 2 — BUNMD Lexis plots 

Lexis plots comparing HMD and BUNMD death coverage.  

```{r}
## Tabulate BUNMD 
bunmd.tabulation <- bunmd %>% 
  filter(death_age %in% 0:100) %>% 
  group_by(Year = dyear, Age = death_age) %>% 
  filter(Year %in% c(1940:2010)) %>% 
  summarize(deaths = n()) 

## HMD tabulation
hmd.tabulation <- hmd %>% 
  filter(Age %in% 0:100) %>% 
  group_by(Year, Age) %>% 
  filter(Year %in% c(1940:2010)) %>% 
  summarize(deaths = sum(Total)) %>% 
  mutate(source = "HMD")

## Prep data for Lexis Function
lexis.data <- bunmd.tabulation %>% 
  inner_join(hmd.tabulation, by = c("Year", "Age" )) %>%
  mutate(coverage = deaths.x/deaths.y)

## Call Lexis Function
lexis_coverage_single <- lexis(data = lexis.data, year_begin = 1940, age_begin = 0, fill_column = coverage, breaks = c(0, 0.25, .5, .75, .9, .95, 5),
                        labels = c( "0 to 25%",
                                    "25% to 50%",
                                    "50% to 75%",
                                    "75% to 90%",
                                    "90% to 95%",
                                    ">95% Death Coverage")) +
  theme(legend.position = "bottom") + 
  theme(axis.title.x = element_text(vjust=-1)) 

## save plot
ggsave(plot = lexis_coverage_single, filename = here("figs/02_bunmd_lexis_coverage.pdf"), width = 6, height = 7)
```

# Figure 3 — Distribution of Age of Death Plot 

Two-panel Plot illustrating effect of truncation on mortality estimation. 

```{r}
## create mortality distribution (untruncated)
trunc_diff <-  tibble(
  pop1  = dgompertz.M(x = 65:110, b = 0.09, M = 80.8),
  pop2  = dgompertz.M(x = 65:110, b = 0.09, M = 85),
  age = 65:110)

## calculate average age of death 
trunc_diff_means <- trunc_diff %>% 
  summarize(pop1_mean = sum(pop1*age)/sum(pop1),
            pop2_mean = sum(pop2*age)/sum(pop2))

## plot untruncated mortality distributions 
truncated_means_plot1 <- trunc_diff %>% 
  ggplot() + 
  geom_line(aes(x = age, y = pop1*100000), color = "red") + 
  geom_line(aes(x = age, y = pop2*100000), color = "blue") + 
  theme_cowplot(rel_large = .75) + 
  labs(x = "Age",
       y = "Deaths",
       title = "Untruncated") + 
  geom_vline(xintercept = trunc_diff_means$pop1_mean, color = "red", linetype = "dashed") + 
  geom_vline(xintercept = trunc_diff_means$pop2_mean, color = "blue", linetype = "dashed") + 
  xlim(65, 110) + 
  ylim(0, 0.04*100000) + 
  annotate("text", x=trunc_diff_means$pop1_mean-1.2, y = 130, 
           label = round(trunc_diff_means$pop1_mean, 1), color = "red") + 
  annotate("text", x=trunc_diff_means$pop2_mean+1.2, y = 130, 
           label = round(trunc_diff_means$pop2_mean, 1), color = "blue") + 
  annotate("text", x = 102.8, y = 3500, label = paste0("Difference = ", round(trunc_diff_means$pop2_mean - trunc_diff_means$pop1_mean, 1), " years"))

## now truncate mortality distribution 
trunc_diff_truncated <- trunc_diff %>% 
  filter(age %in% c(78:95))

## calculate average age of death 
trunc_diff_means_truncated <- trunc_diff_truncated %>% 
  summarize(pop1_mean = sum(pop1*age)/sum(pop1),
            pop2_mean = sum(pop2*age)/sum(pop2))

## plot untruncated mortality distribution 
truncated_means_plot2 <- trunc_diff_truncated %>% 
  ggplot() + 
  geom_line(aes(x = age, y = pop1*100000, color = "Native-born")) + 
  geom_line(aes(x = age, y = pop2*100000,  color = "Foreign-born")) + 
  theme_cowplot(rel_large = .75) + 
  labs(x = "Age",
       y = "Deaths",
      title = "Doubly Truncated") + 
  geom_vline(xintercept = trunc_diff_means_truncated$pop1_mean, color = "red", linetype = "dashed") + 
  geom_vline(xintercept = trunc_diff_means_truncated$pop2_mean, color = "blue", linetype = "dashed") + 
  geom_vline(xintercept = 78, color = "grey") + 
  geom_vline(xintercept = 95, color = "grey") + 
  xlim(65, 110) +
  ylim(0, 0.04*100000) + 
  annotate("text", x = trunc_diff_means_truncated$pop1_mean-2, y = 0.013*100000, 
           label = round(trunc_diff_means_truncated$pop1_mean, 1), color = "red") + 
  annotate("text", x = trunc_diff_means_truncated$pop2_mean+2, y = 0.013*100000, 
           label = round(trunc_diff_means_truncated$pop2_mean, 1), color = "blue") + 
  annotate("text", x = 102.8, y = 3500, 
           label = paste0("Difference = ", round(trunc_diff_means_truncated$pop2_mean - trunc_diff_means_truncated$pop1_mean, 1), " years")) + 
  scale_color_manual(name = "", values = c("Native-born" = "red", "Foreign-born" = "blue")) + 
  theme(legend.position = "bottom")

## truncated means 
truncated_means <- cowplot::plot_grid(truncated_means_plot1, truncated_means_plot2, nrow = 2, labels = "auto")

ggsave(plot = truncated_means, filename = here("figs/03_truncated_means.pdf"), width = 7, height = 6.1)
```

# Figure A-2 — Number apps per person 

```{r}
## number of apps per person
number_apps <- bunmd %>% 
  mutate(number_apps = case_when(
    number_apps >= 3 ~ "3+",
    TRUE ~ as.character(number_apps))) %>% 
    count(number_apps) 

## number of claims per person
number_claims <- bunmd %>% 
  mutate(number_claims = case_when(
    number_claims >= 3 ~ "3+",
    TRUE ~ as.character(number_claims))) %>% 
    count(number_claims) 

## plot claims 
claims_plot <- ggplot(data = number_claims) + 
  geom_col(aes(x = number_claims, y = n), colour="black", fill="grey") + 
  theme(legend.position="bottom") +
  theme_cowplot() + 
  labs(title = "Claims",
       x = "Number of Claim Entries", 
       y = "Count") + 
    scale_y_continuous(labels = scales::comma, limits = c(0, 35000000))
  
## plot apps 
apps_plot <- ggplot(data = number_apps) + 
  geom_col(aes(x = number_apps, y = n), colour="black", fill="grey") + 
  theme(legend.position="bottom") +
  theme_cowplot() + 
  labs(title = "Applications",
       x = "Number of Application Entries", 
       y = "Count") + 
  scale_y_continuous(labels = scales::comma, limits = c(0, 35000000))

## combine plots 
apps_claims_per_person <- cowplot::plot_grid(apps_plot, claims_plot, labels = "auto")

## save plot 
ggsave(plot = apps_claims_per_person, filename = here("figs/a2_apps_claims_per_person.pdf"), width = 10, height = 4)
```

# Figure A-3 — BUNMD Completeness 

```{r}
## recode variables 
bunmd[bunmd == ''| bunmd == ' ' | bunmd == "UNKNOWN" | bunmd == "UN"] <- NA

## calculate percent missing 
bunmd_percent_missing <- bunmd %>% 
   summarise_each(funs(100*mean(!is.na(.)))) %>% 
  mutate(numident_file = "bunmd") %>% 
  pivot_longer(-numident_file)

## recode bunmd completeness
bunmd_completeness <- ggplot(bunmd_percent_missing) + 
  geom_point(aes(x = reorder(name, value), y = value)) + 
  theme_classic() +
  geom_segment(aes(
    x = name,
    xend = name,
    y = 0,
    yend = 100
  ),
  linetype = "dashed",
  size = 0.1
  ) + # Draw dashed lines
  coord_flip() + 
  labs(title = "Berkeley Unified Numident Mortality Database",
       x = "Variable",
       y = "Percent %") +
  ylim(0, 100)

## save ggplot 
ggsave(plot = bunmd_completeness, filename = here("figs/a3_bunmd_completeness.pdf"), width = 10, height = 8)
```