---
title: "Case Studies: Replicate Tables and Figures"
---

Summary: Code to replicate figures and tables in "Berkeley Unified Numident Mortality Database: Public Administrative Records for Individual-Level Mortality Research."

This code replicates Figures 3 and 4. 

## Setup 

Library packages and source functions 

```{r}
## Library packages 
library(tidyverse)
library(data.table)
library(memisc)
library(here)
library(HMDHFDplus)
library(gridExtra)
library(ggpubr)
library(cowplot)
library(gt)
library(fixest)
library(stargazer)
library(choroplethrZip)
library(choroplethr)
library(sf)
library(broom)
library(lubridate)
library(gompertztrunc)

## source functions 
source(here("code/helper_functions.R"))

## Read in BUNMD file
bunmd <- fread(here("data/bunmd_v2.csv"))

## Read in shapefile 
load(here("data/cuyahoga_county_shapefile.rda"))
```

## Merge on zip benefits 

```{r}
## read in SSA benefits data by ZIP Code
oasdi <- fread(here("data/oasdi_zip05_combined_new.csv")) %>% 
  dplyr::select(zip, zip_ben)

## merge on zipcode data 
bunmd[, zip5 := as.numeric(substr(zip_residence, 1, 5))]
bunmd <- merge(bunmd, oasdi[!is.na(zip)], by.x = "zip5", by.y = "zip", all.x = T)
```

# Figure 4 — Immigrant Mortality Advantage 

```{r}
## filter to complete cases
bunmd_cc <- bunmd %>% 
    filter(!is.na(ccweight)) %>% 
    as.data.table()

## set countries 
countries <- c("Canada", "Mexico", "Cuba", "Jamaica", "England", "Ireland", "Greece", "Italy", "Portugal", "Austria", "Philippines", "Japan", "Syria", "Czechoslovakia", "Germany", "Poland", "Yugoslavia", "Russia", "China", "USA")

## restrict to top sending countries  
bunmd_cc <- bunmd_cc %>% 
  mutate(bpl_string = case_when(
    bpl < 1300  ~ "USA",
    bpl %in% 45300:45362 ~ "Germany",
    TRUE ~ bpl_string
  )) %>% 
  filter(bpl_string %in% countries)
```


## Analysis of birth cohorts and years of analysis  

```{r}
## Filter to birth cohorts and years of analysis 
bunmd_analysis <- bunmd_cc %>% 
  filter(byear %in% 1910:1919 &
          dyear %in% 1988:2005 &
          age_first_application < 65) 

## Set reference groups for model
bunmd_analysis <- bunmd_analysis %>% 
  mutate(byear = relevel(as.factor(byear), ref = "1910"),
         country = relevel(as.factor(bpl_string), ref = "USA"))

## Restrict to men 
bunmd_analysis_men <- bunmd_analysis %>% 
  filter(sex == 1)

## Run Gompertz model 
gompertz_results <- gompertz_mle(death_age ~ country, data = bunmd_analysis_men,
                                 left_trunc = 1988, right_trunc = 2005, 
                                 weights = ccweight)

## Convert hazards to life expectancy 
gompertz_results_men <- convert_hazards_to_ex(gompertz_results$results, use_model_estimates = T)

## Plot gompertz results for men 
gompertz_results_men_plot <- gompertz_results_men %>% 
  mutate(parameter = socviz::prefix_strip(parameter, prefixes = "country")) %>% 
  ggplot(aes(y = reorder(parameter, e65), x = e65, xmin = e65_lower, xmax = e65_upper)) + 
  geom_pointrange() + 
  theme_cowplot() + 
    geom_vline(xintercept  = 0, linetype = "dashed") + 
  labs(x = "e65 difference",
       y = "", 
      title =" Foreign−born men, cohorts 1910−1919") + 
  annotate("text", x = 0.1, y = 13, label = "Natives", color = "black", angle = 270) + 
  xlim(-1, 4) 
```


```{r}
## Restrict to women  
bunmd_analysis_women <- bunmd_analysis %>% 
  filter(sex == 2)

## Run Gompertz MLE model 
gompertz_results_women <- gompertz_mle(death_age ~ country, data = bunmd_analysis_women,
                                 left_trunc = 1988, right_trunc = 2005, 
                                 weights = bunmd_analysis_women$ccweight)

## Convert hazards to e65 
gompertz_results_women_e65 <- convert_hazards_to_ex(gompertz_results_women$results, use_model_estimates = T) %>% 
    mutate(estimate = "Gompertz")

## Plot Gompertz results for men 
gompertz_results_women_plot <- gompertz_results_women_e65 %>% 
  mutate(parameter = socviz::prefix_strip(parameter, prefixes = "country")) %>% 
  ggplot(aes(y = reorder(parameter, e65), x = e65, xmin = e65_lower, xmax = e65_upper)) + 
  geom_pointrange() + 
  theme_cowplot() + 
  geom_vline(xintercept  = 0, linetype = "dashed") + 
  labs(x = "e65 difference",
       y = "",
       title =" Foreign−born women, cohorts 1910−1919") + 
  annotate("text", x = 0.1, y = 16, label = "Natives", color = "black", angle = 270) + 
  xlim(-1, 4) 
```

## Combine immigrant mortality advantage plots

```{r}
## combine plots 
gompertz_immigrant_plot <- cowplot::plot_grid(gompertz_results_women_plot, gompertz_results_men_plot, nrow = 2, labels = "auto") 

## save plots 
ggsave(plot = gompertz_immigrant_plot, filename = here("figs/04_immigrant_plot.pdf"), 
       width = 10, height = 10,
       device = cairo_pdf)
```


# Figure 5 — Cuyahoga County Example 

```{r}
## Find cuyahoga county ZIP Codes
data("zip.regions")

## Filter to only include ZIP Codes in Cuyahoga
cuyahoga.zip <- zip.regions %>%
  filter(county.name == "cuyahoga")

## Filter to only include cuyahoga ZIP Codes
cuyahoga.df <- bunmd %>% 
  filter(zip5 %in% cuyahoga.zip$region) %>% 
  as.data.table()

## Restrict to high coverage years
cuyahoga.df.filter <- cuyahoga.df %>% 
  filter(byear %in% 1910:1919 &
           dyear %in% 1988:2005 &
           age_first_application < 65) 

## Prepare variables for regression
cuyahoga.df.filter[, byear := as.factor(byear)]
cuyahoga.df.filter[, byear := relevel(byear, ref = "1910")]

## Prepare variables for regression
cuyahoga.df.filter[, zip5 := as.factor(zip5)]
cuyahoga.df.filter[, zip5 := relevel(zip5, ref = "44101")]

## Gompertz model 
cuyahoga.model_gompertz <- gompertztrunc::gompertz_mle(death_age ~ zip5 + as.factor(sex), right_trunc = 2005, left_trunc = 1988,
                     data = cuyahoga.df.filter,
                     weights = ccweight)

## convert hazards to life expectancuy 
gompertz_cuyahoga <- gompertztrunc::convert_hazards_to_ex(cuyahoga.model_gompertz$results)

## clean model output 
gompertz_cuyahoga <- gompertz_cuyahoga %>% 
  mutate(region = substr(parameter, 5, 9)) %>% 
  mutate(value = e65) %>% 
  mutate(value = round(value, 1))

## Create choro plot
ohio_fips <-  39035
choro = ZipChoropleth$new(rev(gompertz_cuyahoga))
choro$title = ""
choro$set_zoom_zip(state_zoom = NULL, county_zoom = ohio_fips, msa_zoom = NULL, zip_zoom = NULL)

## Manually set scale 
scale_colour_manual(values = c("d73027", "fc8d59", "fee090", "ffffbf", "e0f3f8", "91bfdb", "4575b4"))

## Plot Cleveland
cleveland_plot <- choro$render() +
  theme(text=element_text(size=25)) + 
  scale_colour_manual(
    values = c("#d73027", "#fc8d59", "#fee090", "#ffffbf", "#e0f3f8", "#91bfdb", "#4575b4"),
    aesthetics = c("colour", "fill"),
    name = "Difference in e65"
  ) + 
  geom_sf(data = st_union(cuyahoga_county_shapefile),
       inherit.aes = FALSE, fill = NA, size = 1.2, color = "Black") +
  annotate("label", x=-81.67908, y=41.48074, label="Cleveland", color = "Black", size = 8)
```

## Cuyahoga county ZIP benefits 

```{r}
## Calculate benefits per ZIP Code
zipben_cuyahoga <- cuyahoga.df %>% 
  group_by(zip5) %>% 
  summarize(zip_ben = round(1000*mean(zip_ben))) %>% 
  dplyr::select(region = zip5, value = zip_ben) %>% 
  distinct() %>% 
  mutate(region = as.character(region))

## Ohio Plots 
ohio_fips <-  39035
choro = ZipChoropleth$new(rev(zipben_cuyahoga))
choro$title = ""
choro$set_zoom_zip(state_zoom = NULL, county_zoom = ohio_fips, msa_zoom = NULL, zip_zoom = NULL)

## Plot Cleveland
cleveland_plot_zipben <- choro$render() +
  theme(text=element_text(size=25)) + 
   # viridis::scale_fill_viridis(discrete = "true", direction = 1, name = "Avg. SS Benefits") + #  
  scale_colour_manual(
    values = c("#d73027", "#fc8d59", "#fee090", "#ffffbf", "#e0f3f8", "#91bfdb", "#4575b4"),
    aesthetics = c("colour", "fill"),
    name = "Avg. SS Benefits"
  ) + 
  geom_sf(data = st_union(cuyahoga_county_shapefile),
       inherit.aes = FALSE, fill = NA, size = 1.2, color = "Black") +
  annotate("label", x=-81.67908, y=41.48074, label="Cleveland", color = "Black", size = 8) 
```

## Save plots and write out 

```{r}
## combine cleveland plots into a single plot 
cleveland_plot_combined <- cowplot::plot_grid(cleveland_plot, cleveland_plot_zipben, nrow = 2, labels = "auto", label_size = 26)

## Save plot 
ggsave(plot = cleveland_plot_combined, filename = here("figs/05_cleveland_plot_combined.pdf"), width = 14, height = 18)
```



