---
title: "Generate BUNMD Weights"
author: Casey Breen
date: 6/19/2022
---

## Summary

This notebook contains code to generate a set of person-level weights for the BUNMD for different custom sub-samples. The set of person-level weights will make the weighted count of deaths in the BUNMD match aggregate population totals from the Human Mortality Database (HMD). To construct the weights, we spit the BUNMD into cells cross-classified by year of birth, year of death, age at death, and sex. We assigned every person in a given cell a weight equal to the ratio of count of deaths in the HMD to count of deaths in the BUNMD sample: 

\begin{equation}
  W_{j}  = \frac{\text{HMD deaths in cell j}}{\text{BUNMD deaths in cell j}}.
\end{equation}

Note: This weighting strategy does not take geography into account and assumes uniform coverage for other covariates, such as race, socioeconomic status, and household size. 

## Functions to create weights for custom sub-samples

By default, this function constructs weights for those dying between 1988 and 2005, born between 1895 and 1940, dying between ages of 65 and 100, with non-missing values for race (`race_last`), sex (`sex`), and birthplace (`bpl`). However, the arguments of the `create_weights_bunmd_complete()` function can be changed. For instance, if a user was only using birthplace in their analysis, they could only require the birthplace (`bpl`) variable to be non-missing.  

To run this script, will need to download a copy of the `USA.Deaths_lexis.txt` file from HMD (https://www.mortality.org/Country/Country?cntr=USA). 

## Set-up 

```{r}
## library packages 
library(tidyverse)       ## data manipulation and visualization 
library(data.table)      ## data.table 
library(here)            ## creates paths relative to the top-level directory

## Read in bunmd and set blank values ("") to NA 
## Update path 
bunmd <- fread(here("data/bunmd_v2.csv"), na.strings = c(NA_character_, ""))

## Lexis Triangles (Deaths) from BUNMD
## Update path
hmd_deaths <- fread(here("code/weights/Deaths_lexis.txt")) %>%
    mutate(linking_key = paste(Year, Cohort, Age, sep = "_" ))
```

## Weighting Function 

```{r}
#' Create weights for BUNMD
#' This function will weight the BUNMD file up to HMD lexis triangles
create_weights_bunmd_complete <- function(bunmd.file, dyears = c(1988:2005), cohorts = c(1895:1940), death_ages = c(65:100), complete_cases_vars = c("race_last", "bpl", "sex"), hmd_deaths = hmd_deaths) {

  ## filter to analytic sample of interest 
  high_coverage_complete_sample <- bunmd.file %>%
    filter(dyear %in% dyears) %>%
    filter(byear %in% cohorts) %>%
    filter(death_age %in% death_ages) %>%
    filter(!is.na(sex)) %>%
    drop_na(complete_cases_vars)

  ## tabulate number of deaths in cells cross-classified by age of death, year of death, year of birth, and sex (lexis triangles)
  numdeath_aggregate_counts <- high_coverage_complete_sample %>%
    group_by(death_age, dyear, byear, sex) %>%
    tally() %>%
    mutate(linking_key = paste(dyear, byear, death_age, sep = "_")) %>%
    ungroup(dyear, death_age)

  ## create inverse probability weights
  death_weights <- numdeath_aggregate_counts %>%
    inner_join(hmd_deaths, by = "linking_key") %>%
    mutate(proportion_matched = case_when(
      sex == 1 ~ n/Male,
      sex == 2 ~ n/Female)) %>%
    group_by(dyear, byear, sex, death_age) %>%
    summarize(inclusion_prob = mean(proportion_matched), Male = mean(Male), Female = mean(Female), Total = mean(Total))

  ## create death weights
  death_weights_for_link <-  death_weights %>%
    mutate(linking_key = paste(dyear, byear, death_age, sex, sep = "_")) %>%
    ungroup(dyear, death_age, sex) %>%
    select(inclusion_prob, linking_key) %>%
    mutate(weight_alt = 1/inclusion_prob) %>%
    select(-inclusion_prob)

  ## create linking key
  high_coverage_complete_sample <- high_coverage_complete_sample %>%
    mutate(linking_key = paste(dyear, byear, death_age, sex, sep = "_")) %>%
    select(-byear)

  ## link weights by ssn
  weights.df <- high_coverage_complete_sample %>%
    left_join(death_weights_for_link, by = "linking_key") %>%
    select(ssn, weight_alt)

  ## add weights back to original sample
  bunmd.file <- bunmd.file %>%
    left_join(weights.df, by = "ssn")

  return(bunmd.file)
}
```

## Apply weighting function 

The function below will create a new weight variable `weight_alt.` The default options below will construct new variable `weight_alt` identical to the complete case weight (`ccweight`), but the arguments of the function can be changed to create custom weights. 

```{r}
## example construct weights require bpl, race_last, and sex variables to be non-missing 
## new weight variable = weight_alt
bunmd <-  create_weights_bunmd_complete(bunmd.file = bunmd, dyears = c(1988:2005), cohorts = c(1895:1940), complete_cases_vars = c("race_last", "bpl", "sex"), death_ages = c(65:100), hmd_deaths = hmd_deaths)
```



